---
title: "motMis - study 4a"
output: html_document
---
```{r install/load packages,include=FALSE}

#Checks if required packages are installed. Installs them if not.
  # list_pckgs <- c("tidyverse", "psych","lsr","mblm","broom", "effsize","matrixStats","devtools","yarrr","lmtest","lubridate","rstudioapi","apaTables","broman") # list of required packages
  # new_pckgs <- list_pckgs[!(list_pckgs %in% installed.packages()[,"Package"])] # list of required packages that are not install on this machine
  # if(length(new_pckgs)) install.packages(new_pckgs) # installs them

#Load libraries
  library("tidyverse") # plotting, manipulating data etc
  library("psych") # for trait measures
  library("lsr") # for computing effect sizes
  library("mblm") #non-parametric linear regression
  library("broom") # clean outputs
  library("effsize") #non-parametric effect sizes
  library("matrixStats") #calculation SDs for multiple cols
  library("devtools") #pirateplots
  library("yarrr")#pirateplots
  library("lmtest")
  library("lubridate")
  library("rstudioapi") # load rstudio api
  library("apaTables") # for regression tables
  library("broman") # flexibly display p-values with Rmd function
  library("MBESS") # CIs on Rsquared in tables
  
#Set working directory to the location of this R file (data should also be saved here)
  # curr_dir <- getActiveDocumentContext()$path # get path to currently open R file
  # setwd(dirname(curr_dir)) # set the working directory to curr_dirr
  # #print(getwd()) # print directory
  # rm("curr_dir","new_pckgs","list_pckgs")

```

```{r non-parametric stats functions,include=FALSE}

#wilcoxon test - paired
  w.test = function(x,y){
  t = wilcox.test(x,y, paired=TRUE) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x,y) 
    da = cohen.d(x,y)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = da$estimate #add parametric ES
      tbl$cliff.d = d$estimate #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }

#wilcoxon test - one-sample
  w.test_0 = function(x){
  t = wilcox.test(x,mu=0) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x,0) 
    da = cohensD(x,mu=0)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = da #add parametric ES
      tbl$cliff.d = d$estimate #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }

#mann-whitney test 
  mw.test = function(x,y){
  t = wilcox.test(x~y) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x~y) 
    da = cohen.d(x~y)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = abs(da$estimate) #add parametric ES
      tbl$cliff.d = abs(d$estimate) #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }
```

```{r read in data set, include=FALSE}

#read data
  df4a = read.csv("motMis_s4a_data.csv") #note: N = 1154, but only N = 1152 had complete data (see note below)
   
#note on missing data 
  which(is.na(df4a$fairQ_1)) # 2 Ps missing key measure of personal standards. These 2 Ps (IDs 770 & 1122 in raw data) were accidentally used as pairs in Study 4b. As reported in the MS, these Ps and their pairs in S4B were excluded from analyses. See r exclusions chunk.
  
  #df4a = df4a[!is.na(df4a$fairQ_1),] # code to remove two participants with missing data for accurately reporting gender and age

#gender and age (see note above)
  describe(df4a$age)$mean
  table(df4a$gender)
 
```

```{r wrangling, include=FALSE}

#rename vars
  colnames(df4a)[colnames(df4a)=="fairQ_1"] <- "fairQ"
  colnames(df4a)[colnames(df4a)=="dg1_1"] <- "dg1"
  colnames(df4a)[colnames(df4a)=="dg2_1"] <- "dg2"
  colnames(df4a)[colnames(df4a)=="dg3_1"] <- "dg3"
  colnames(df4a)[colnames(df4a)=="dg4_1"] <- "dg4"
  colnames(df4a)[colnames(df4a)=="dg5_1"] <- "dg5"
  colnames(df4a)[colnames(df4a)=="dg1_rt_Page.Submit"] <- "dg1_rt"
  colnames(df4a)[colnames(df4a)=="dg2_rt_Page.Submit"] <- "dg2_rt"
  colnames(df4a)[colnames(df4a)=="dg3_rt_Page.Submit"] <- "dg3_rt"
  colnames(df4a)[colnames(df4a)=="dg4_rt_Page.Submit"] <- "dg4_rt"
  colnames(df4a)[colnames(df4a)=="dg5_rt_Page.Submit"] <- "dg5_rt"
  colnames(df4a)[colnames(df4a)=="memQ_4"] <- "Guess"

#trim vars
  df4a$responsible = trimws(df4a$responsible, which ="left") # remove extra spaces in variable
  df4a$responsible = as.numeric(df4a$responsible)
  df4a$conf = trimws(df4a$conf, which ="left") # remove extra spaces in variable
  df4a$conf = as.numeric(df4a$conf)

```

```{r compute variables, include=FALSE}

#compute the true average % transferred + average RT
  df4a <- df4a %>% mutate(DGAverage=(dg1+dg2+dg3+dg4+dg5)/5, AveRT=(dg1_rt+dg2_rt+dg3_rt+dg4_rt+dg5_rt)/5) 
    df4a$DGAverage = as.numeric(df4a$DGAverage)

#compute the % difference between recalled transfer vs. actual transfer
  df4a$GuessVsAve = (df4a$Guess - df4a$DGAverage)

#compute absolute % difference between recalled vs. actual transfer
  df4a$GuessVsAveA = abs(df4a$GuessVsAve)
  
#compute standard deviation for the five transfer decisions for each subject
  df4a_sd <- c('dg1', 'dg2', 'dg3', 'dg4', 'dg5')
    df4a <- df4a %>% mutate(DG.stdev=rowSds(as.matrix(.[df4a_sd])))

#compute difference score between fairness vs actual transfer 
  df4a$FairVsActual = (df4a$fairQ - df4a$DGAverage)

#compute discomfort_t1
  keys.list <- list(discomfort_t1=c("discomfort1_1","discomfort1_2","discomfort1_3"))
  discomfort_t1 <- scoreItems(keys.list,df4a)
  discomfort_t1
  summary(discomfort_t1)
  describe(discomfort_t1$scores)
  df4a$discomfort_t1 <- discomfort_t1$scores #add discomfort composite measure to dataset 
    
#compute discomfort_t2
  keys.list <- list(discomfort_t2=c("discomfort3_1","discomfort3_2","discomfort3_3"))
  discomfort_t2 <- scoreItems(keys.list,df4a)
  summary(discomfort_t2)
  describe(discomfort_t2$scores)
  df4a$discomfort_t2 <- discomfort_t2$scores #add discomfort composite measure to dataset
  
#compute dissonance
  df4a$diss = as.numeric(df4a$discomfort_t2 - df4a$discomfort_t1) 
  df4a$diss
  
#compute posA_t1
  keys.list <- list(posA_t1=c("discomfort1_4","discomfort1_5","discomfort1_6","discomfort1_7","discomfort1_8"))
  posA_t1 <- scoreItems(keys.list,df4a)
  posA_t1
  summary(posA_t1)
  describe(posA_t1$scores)
  df4a$posA_t1 <- posA_t1$scores #add discomfort composite measure to dataset 
  
#compute posA_t2
  keys.list <- list(posA_t2=c("discomfort3_4","discomfort3_5","discomfort3_6","discomfort3_7","discomfort3_8"))
  posA_t2 <- scoreItems(keys.list,df4a)
  posA_t2
  summary(posA_t2)
  describe(posA_t2$scores)
  df4a$posA_t2 <- posA_t2$scores #add discomfort composite measure to dataset
  
#compute posAdiff
  df4a$posAdiff = as.numeric(df4a$posA_t2 - df4a$posA_t1) 
  df4a$posAdiff
  
rm("keys.list")

```

```{r reformat vars,include=FALSE} 

  df4a$Guess = as.numeric(df4a$Guess)
  df4a$DGAverage = as.numeric(df4a$DGAverage)
  df4a$AveRT = as.numeric(df4a$AveRT)
  df4a$DG.stdev = as.numeric(df4a$DG.stdev)
  df4a$FairVsActual = as.numeric(df4a$FairVsActual)
  df4a$discomfort_t1 = as.numeric(df4a$discomfort_t1)
  df4a$discomfort_t2 = as.numeric(df4a$discomfort_t2)
  df4a$diss = as.numeric(df4a$diss)
  df4a$posA_t1 = as.numeric(df4a$posA_t1)
  df4a$posA_t2 = as.numeric(df4a$posA_t2)
  df4a$posAdiff = as.numeric(df4a$posAdiff)
  df4a$conf = as.numeric(df4a$conf)
  df4a$studytime <- ms(df4a$studytime) # using lubridate library to convert time
  df4a$studytime <- as.numeric(df4a$studytime) #minutes converted to numeric
  
```

```{r exclusions,include=FALSE} 

#EXCLUSIONS

  #1 Failed to follow instructions: bottest = 0 (failed bottest)
    # count(df4a$bottest_filter == 0)
    df4a <- subset(df4a, bottest_filter == 1)

  #2 Failed comprehension: deception1 = 0 (instructions NOT clear about money)
    #count(df4a$deception1 == 0)
    df4a <- subset(df4a, deception1 == 1)

  #3 Failed comprehension: deception2 = 0 (intructions NOT clear about other participant)
    #count(df4a$deception2 == 0)
    df4a <- subset(df4a, deception2 == 1)

  #4 Reported skepticism about the study: feedbackexcl = 0 (did NOT believe in P2)
    count(df4a$feedbackexcl <= 0)
    df4a <- subset(df4a, feedbackexcl > 0)

  #5 Outliers on our key measure: GuessVsAve
    #ggplot(df4a, aes("",GuessVsAve))+geom_boxplot()+theme_classic() # before
    stats = psych::describe(df4a$GuessVsAve) #exclude extreme outliers (4 SDs from the mean).
    high_cutoff = stats$mean + (stats$sd*4)
    low_cutoff = stats$mean - (stats$sd*4)

    #SD outliers for memory errors
    count(df4a$GuessVsAve >= high_cutoff)
    df4a <- subset(df4a, GuessVsAve < high_cutoff)

    count(df4a$GuessVsAve <= low_cutoff)
    df4a <- subset(df4a, GuessVsAve > low_cutoff)
  #ggplot(df4a, aes("",GuessVsAve))+geom_boxplot()+theme_classic() # after

rm("stats","high_cutoff","low_cutoff")


#SUPPLEMENTAL EXCLUSIONS:

  #Exclude people who never transferred a positive amount
     #df4a <- subset(df4a, DGAverage > 0)
     #df4a_0 <- subset(df4a, DGAverage == 0)

  #Static deciders: Exclude people who always gave the same amount
    #df4a <- subset(df4a, DG.stdev > 0)

  #Responsibility: Exclude based on perceived responsibility
    #df4a <- subset(df4a, responsible == 1)
    #df4a <- subset(df4a, responsible > 1)

```

```{r export choices for s4b pairs (pre-study), echo=FALSE,include=FALSE}

  # this chunk was used to export (free) choice sets for the forced choice study

  # a=df4a$dg1 
  # b=df4a$dg2
  # c=df4a$dg3 
  # d=df4a$dg4 
  # e=df4a$dg5 
  # decisions = cbind(a,b,c,d,e)
  # decisions = as.data.frame(decisions)

  # write.csv(decisions, file = "motMis_s4a_decisions.csv") # N = 711, decisions used for forced choice study (including 2 Ps with missing data who were included due to an oversight)

  df4a$mergeID <- seq.int(nrow(df4a)) # creates an ID variable that replicates the IDs exported with free-choice decision sets

```

```{r exclude Ps with missing data (post-study), echo=FALSE,include=FALSE}

  # follow-up note: as seen below, two Ps did not respond to our crucial fairQ DV, and thus should have been excluded
    which(is.na(df4a$fairQ)) # rows 476 & 692 (after above exclusions)

  # note, exclude s4b Ps paired with rows 476 & 692 (IDs 770 & 1122), which have missing data. Also, row 89 in topup.
    subset(df4a[476,], select=c(fairQ, dg1, dg2, dg3, dg4, dg5))
    subset(df4a, id == 770, select=c(fairQ, dg1, dg2, dg3, dg4, dg5))
    subset(df4a[692,], select=c(fairQ, dg1, dg2, dg3, dg4, dg5)) 
    subset(df4a, id == 1122, select=c(fairQ, dg1, dg2, dg3, dg4, dg5))
    
  # code below removes the two participants with missing data from s4a (N=709)
    df4a = df4a[!is.na(df4a$fairQ),] 

```

```{r import memories from s4b pairs (post-study), echo=FALSE,include=FALSE}

  #import forced choice data (run s4b.Rmd to produce .csv)
    df4b = read.csv("motMis_s4b_mems.csv") # read in memories from forced choice study

  # #pull forced choices to ensure they match free choices
  #   a=df4b$dg1
  #   b=df4b$dg2
  #   c=df4b$dg3
  #   d=df4b$dg4
  #   e=df4b$dg5
  # 
  #   forcedD = cbind(a,b,c,d,e)
  #   forcedD = as.data.frame(forcedD)
  #   forcedD$mergeID = df4b$id.1
  #   forcedD$recall <- df4b$Guess
  #   forcedD$recallDiff <- df4b$GuessVsAve
  # 
  #   merged_dg = merge(df4a, forcedD, by = "mergeID")
  #   df4a=merged_dg
  # 
  # #test if choices match
  #   df4a$a == df4a$dg1 # test1
  #   df4a$e == df4a$dg5 # test2
  # 
  # #redefine recall as forced choice participants' responses
  #   df4a$Guess = forcedD$recall
  #   df4a$GuessVsAve = forcedD$recallDiff

  #knit to run analyses

```

```{r grouping, include=FALSE, echo=FALSE}

#GIVERS vs. NON-GIVERS: did they give a positive amount?

  df4a$giver = as.numeric(df4a$DGAverage > 0)


#SELFISH vs. GENEROUS: median split participants based on their average transfers

  m = median(df4a$DGAverage,na.rm = TRUE) 
  df4a_selfish <- subset(df4a, DGAverage < m) 
  df4a_generous <- subset(df4a, DGAverage >= m) 
  
  df4a$DGAverage.split = NA 
  df4a$DGAverage.split[df4a$DGAverage < m] <- "selfish" 
  df4a$DGAverage.split[df4a$DGAverage >= m] <- "generous" 
  df4a$DGAverage.split <- as.factor(df4a$DGAverage.split)
    
  
#NORM VIOLATORS vs. NORM UPHOLDERS: split by whether people upheld or violated their own fairness norms
  
  a=df4a$dg1 < df4a$fairQ
  b=df4a$dg2 < df4a$fairQ
  c=df4a$dg3 < df4a$fairQ
  d=df4a$dg4 < df4a$fairQ
  e=df4a$dg5 < df4a$fairQ
  df4a$violations <- a+b+c+d+e
  
  df4a_violators <- subset(df4a, DGAverage < fairQ) 
  df4a_upholders <- subset(df4a, DGAverage > fairQ) 

  df4a$Fair.split = NA 
  df4a$Fair.split[df4a$DGAverage < df4a$fairQ] <- "violators" 
  df4a$Fair.split[df4a$DGAverage > df4a$fairQ] <- "upholders"
  df4a$Fair.split <- as.factor(df4a$Fair.split)
  
    
#UNGENEROUS vs. GENEROUS norm upholders: split on generosity within fairness upholders
  
  m_fair = median(df4a_upholders$DGAverage,na.rm = TRUE) 
  df4a_fair_selfish <- subset(df4a_upholders, DGAverage < m_fair)
  df4a_fair_generous <- subset(df4a_upholders, DGAverage > m_fair)
  
  df4a_upholders$DGAverage.split = NA 
  df4a_upholders$DGAverage.split[df4a_upholders$DGAverage < m_fair] <- "selfish"
  df4a_upholders$DGAverage.split[df4a_upholders$DGAverage > m_fair] <- "generous"  
  df4a_upholders$DGAverage.split <- as.factor(df4a_upholders$DGAverage.split)


#MOTIVATED MEMORY: create variables capturing positive (motivated) or negative (nonmotivated) memory errors
  
  df4a$motmem = NA
  df4a$motmem[df4a$GuessVsAve < 0] = 0
  df4a$motmem[df4a$GuessVsAve > 0] = 1
  df4a$motmem = as.factor(df4a$motmem)
  
  df4a_upholders$motmem = NA
  df4a_upholders$motmem[df4a_upholders$GuessVsAve < 0] = 0
  df4a_upholders$motmem[df4a_upholders$GuessVsAve > 0] = 1
  df4a_upholders$motmem = as.factor(df4a_upholders$motmem)
    table(df4a_upholders$motmem)
  
  df4a_violators$motmem = NA
  df4a_violators$motmem[df4a_violators$GuessVsAve < 0] = 0
  df4a_violators$motmem[df4a_violators$GuessVsAve > 0] = 1
  #df4a_violators$motmem[df4a_violators$GuessVsAve == 0] = 2
  df4a_violators$motmem = as.factor(df4a_violators$motmem)
    table(df4a_violators$motmem)
   

    
###########################################  
### test of norm-directed memory errors ###
     
d5=table(df4a_violators$motmem) #violators TWICE AS LIKELY to misremember in one direction
binom.test(d5, p = 0.5, conf.level = 0.95)
1-(binom.test(d5, p = 0.5, conf.level = 0.95)$estimate)
1-(binom.test(d5, p = 0.5, conf.level = 0.95)$conf.int)
  d5

d6=table(df4a_upholders$motmem) #exceeders AT CHANCE (should be remembering less)
binom.test(d6, p = 0.5, conf.level = 0.95)
1-(binom.test(d6, p = 0.5, conf.level = 0.95)$estimate)
1-(binom.test(d6, p = 0.5, conf.level = 0.95)$conf.int)
  d6
    
    
```

```{r sanity checks, warning=FALSE, message=FALSE, echo=FALSE,include=FALSE}

#[a] are memory errors non-normally distributed, visually? 
  ggplot(df4a,aes(GuessVsAve))+geom_histogram(binwidth=2)

#[b] are memory errors non-normally distributed, statistically? 
  shapiro.test(df4a$GuessVsAve) 

#[c] can memory errors be explained by noise in DG offers?
  cor.test(df4a$DG.stdev,df4a$GuessVsAveA, method="spearman") 

#[d] can memory errors be explained by RT (attention) to DG offers?
  cor.test(df4a$AveRT,df4a$GuessVsAveA, method="spearman") 

#[e] can memory errors be explained by non-giving?
  cor.test(df4a$giver,df4a$GuessVsAveA, method="spearman") 

#[e] can memory errors be explained by how long people took to complete the study?
  cor.test(as.numeric(df4a$studytime),df4a$GuessVsAveA,method="spearman")

#[f] can memory errors be explained by numeracy?
  cor.test(df4a$Numeracy,df4a$GuessVsAveA, method="spearman") 
  
#[g] were participants responsible for their actions?
  mean(df4a$responsible)
  sd(df4a$responsible)
  count(df4a$responsible==7)
  nrow(df4a)
  count(df4a$responsible==7)/nrow(df4a)
  
  
# group differences
  mw.test(df4a$DG.stdev,df4a$Fair.split)
  mw.test(df4a$AveRT,df4a$Fair.split)
  mw.test(df4a$giver,df4a$Fair.split)
  mw.test(df4a$Numeracy,df4a$Fair.split)

  
```


```{r tests,echo=FALSE,warning=FALSE,include=FALSE}

#[Check normality]
  norm = shapiro.test(df4a$GuessVsAve)

#[Descriptives]
  df4a_d = describe(df4a$DGAverage)
  df4a_v_d = describe(df4a_violators$DGAverage)
  df4a_u_d = describe(df4a_upholders$DGAverage)
  df4a_r = describe(df4a$Guess)
  df4a_v_r = describe(df4a_violators$Guess)
  df4a_u_r = describe(df4a_upholders$Guess)

#[1] Motivated misremembering (main effect)
  t1 = w.test_0(df4a$GuessVsAve)

#[2] Motivated misremembering: VIOLATOR group
  t2 = w.test_0(df4a_violators$GuessVsAve)

#[3] Motivated misremembering: UPHOLDER group
  t3 = w.test_0(df4a_upholders$GuessVsAve)

#[4] Motivated misremembering: group difference
  t4 = mw.test(df4a$GuessVsAve,df4a$Fair.split)
  
#[5] Memory accuracy: group difference
  t5 = mw.test(df4a$GuessVsAveA,df4a$Fair.split)

##################################################
  
#[Descriptives]
  df4a_fu_d = describe(df4a_fair_selfish$DGAverage)
  df4a_fg_d = describe(df4a_fair_generous$DGAverage)
  
#[6] Motivated misremembering: fair UNGENEROUS group
  t6 = w.test_0(df4a_fair_selfish$GuessVsAve)

#[7] Motivated misremembering: fair GENEROUS group
  t7 = w.test_0(df4a_fair_generous$GuessVsAve)
 
#[8] Motivated misremembering: group difference
  t8 = mw.test(df4a_upholders$GuessVsAve,df4a_upholders$DGAverage.split)

##################################################

#[9] Model predicting recalled generosity from actual geneorsity (controlling for decision speed + variance + non-giving + numeracy)
    df4a_lm = select(df4a,Guess,DGAverage,AveRT,DG.stdev,giver,FairVsActual,Numeracy)
    df4a_lm = na.omit(df4a_lm)

    t9 <- lm(Guess ~ DGAverage + AveRT + DG.stdev + giver + Numeracy , data=df4a_lm)
    tidy(t9)
    t9a = AIC(t9)
    t9b = BIC(t9)
    t9l = logLik(t9)
      t9l
      
#[10] Model predicting recalled generosity from actual geneorsity fairness deviations (controlling for decision speed + variance + non-giving + numeracy)
    t10 <- lm(Guess ~ DGAverage + AveRT + DG.stdev + giver + Numeracy + FairVsActual, data=df4a_lm)
    tidy(t10)
    t10a = AIC(t10)
    t10b = BIC(t10)
    t10l = logLik(t10)
      t10l

#[11] Comparison of models' goodness of fit
    t11 = lrtest(t9,t10)
    t11

##################################################

#[12] Confidence: Correlations
  t12 = cor.test(df4a$conf,df4a$GuessVsAveA, method="spearman") 
  
#[13] Confidence: Group difference
  t13 = mw.test(df4a$conf,df4a$Fair.split)

#[14] Discomfort and positive affect
  t14a = mw.test(df4a$discomfort_t2,df4a$Fair.split)
  t14b = mw.test(df4a$diss,df4a$Fair.split)
  t14c = mw.test(df4a$posA_t2,df4a$Fair.split)
  t14d = mw.test(df4a$posAdiff,df4a$Fair.split)

#[15] Discomfort benefits of memory errors: Upholders & Violators
  t15a = kruskal.test(diss ~ Fair.split, data = df4a)
  t15b = kruskal.test(diss ~ motmem, data = df4a)
  int = interaction(df4a$Fair.split, df4a$motmem)
  t15c = kruskal.test(diss ~ int, data = df4a)  
  t15d = mw.test(df4a_upholders$diss,df4a_upholders$motmem)
  t15e = mw.test(df4a_violators$diss,df4a_violators$motmem)

```

```{r knit functions,echo=FALSE}

#formatting p-values for APA
format_pval <- function(x){
  if (x < .001) return(paste('<', '.001'))
  if (x > .1) return(paste('=', myround(x, 2)))
  paste('=', myround(x, 3))   # 3 = no. of digits to round p value to if p < .001 
}

```
#####Check normality

(*W* = `r myround(norm$s,2)`, *p* `r format_pval(norm$p)`)

#####Descriptives

N = `r round(df4a_d$n,1)`, %given (M = `r myround(df4a_d$mean,2)`, SD = `r myround(df4a_d$sd,2)`), %recalled (M = `r myround(df4a_r$mean,2)`, SD =`r myround(df4a_r$sd,2)`)  
Violators N = `r round(df4a_v_d$n,1)`, %given (M = `r myround(df4a_v_d$mean,2)`, SD = `r myround(df4a_v_d$sd,2)`), %recalled (M = `r myround(df4a_v_r$mean,2)`, SD =`r myround(df4a_v_r$sd,2)`)  
Upholders N = `r round(df4a_u_d$n,1)`, %given (M = `r myround(df4a_u_d$mean,2)`, SD = `r myround(df4a_u_d$sd,2)`), %recalled (M = `r myround(df4a_u_r$mean,2)`, SD =`r myround(df4a_u_r$sd,2)`)  

#####[1] Motivated misremembering (main effect)  

[1] (*V* = `r format(t1$s,scientific=FALSE)`, *p* `r format_pval(t1$p)`, *d* = `r myround(t1$coh,2)`, δ = `r myround(t1$cl,2)`)

#####[2] Motivated misremembering: VIOLATORS group 

[2] (*V* = `r format(t2$s,scientific=FALSE)`, *p* `r format_pval(t2$p)`, *d* = `r myround(t2$coh,2)`, δ = `r myround(t2$cl,2)`)

#####[3] Motivated misremembering: UPHOLDERS group 

[3] (*V* = `r format(t3$s,scientific=FALSE)`, *p* `r format_pval(t3$p)`, *d* = `r myround(t3$coh,2)`, δ = `r myround(t3$cl,2)`)

#####[4] Motivated misremembering: group difference  

[4] (*W* = `r format(t4$s,scientific=FALSE)`, *p* `r format_pval(t4$p)`, *d* = `r myround(t4$coh,2)`, δ = `r myround(t4$cl,2)`)

#####[5] Memory accuracy: group difference  

[5] (*W* = `r format(t5$s,scientific=FALSE)`, *p* `r format_pval(t5$p)`, *d* = `r myround(t5$coh,2)`, δ = `r myround(t5$cl,2)`)

#####Descriptives

N = `r round(df4a_d$n,1)`, M = `r myround(df4a_d$mean,1)`   
Fair ungenerous N = `r round(df4a_fu_d$n,1)`, M = `r myround(df4a_fu_d$mean,1)`   
Fair generous N = `r round(df4a_fg_d$n,1)`, M = `r myround(df4a_fg_d$mean,1)`

#####[6] Motivated misremembering: FAIR ungenerous group 

[6] (*V* = `r format(t6$s,scientific=FALSE)`, *p* `r format_pval(t6$p)`, *d* = `r myround(t6$coh,2)`, δ = `r myround(t6$cl,2)`)

#####[7] Motivated misremembering: FAIR generous group 

[7] (*V* = `r format(t7$s,scientific=FALSE)`, *p* `r format_pval(t7$p)`, *d* = `r myround(t7$coh,2)`, δ = `r myround(t7$cl,2)`)

#####[8] Motivated misremembering: group difference  

[8] (*W* = `r format(t8$s,scientific=FALSE)`, *p* `r format_pval(t8$p)`, *d* = `r myround(t8$coh,2)`, δ = `r myround(t8$cl,2)`)

#####[9-11] Model comparison: recall from actual generosity (i) or actual generosity + fairness deviations (ii)

[9-11] (ΔAIC = `r myround(t9a-t10a,2)`, ΔBIC = `r myround(t9b-t10b,2)`, χ2 (`r (t11$D[2])`) = `r myround(t11$Ch[2],2)`, *p* `r format_pval(t11$P[2])`)

#####[12] Confidence: Correlations

[12]  Unsigned memory errors (*r~s~* = `r myround(t12$e,2)`, *p* `r format_pval(t12$p.v)`)

#####[13] Confidence: Group difference
[13] (*W* = `r format(t13$s,scientific=FALSE)`, *p* `r format_pval(t13$p)`, *d* = `r myround(t13$coh,2)`, δ = `r myround(t13$cl,2)`)

#####[14] Moral and emotional benefits
[14a] Discomfort (*W* = `r format(t14a$s,scientific=FALSE)`, *p* `r format_pval(t14a$p)`, *d* = `r myround(t14a$coh,2)`, δ = `r myround(t14a$cl,2)`)   
[14b] Discomfort Diff (*W* = `r format(t14b$s,scientific=FALSE)`, *p* `r format_pval(t14b$p)`, *d* = `r myround(t14b$coh,2)`, δ = `r myround(t14b$cl,2)`)  
[14c] Positive Affect (*W* = `r format(t14c$s,scientific=FALSE)`, *p* `r format_pval(t14c$p)`, *d* = `r myround(t14c$coh,2)`, δ = `r myround(t14c$cl,2)`)   
[14d] Positive Affect Diff (*W* = `r format(t14d$s,scientific=FALSE)`, *p* `r format_pval(t14d$p)`, *d* = `r myround(t14d$coh,2)`, δ = `r myround(t14d$cl,2)`)

#####[15] Discomfort Diff benefits of memory errors: Upholders & Violators

[15a] 2x2 Main effect: Fairness Standards (*H~kruskal.wallis~*(`r (t15a$pa)`) = `r myround(t15a$s,2)`, *p* `r format_pval(t15a$p.v)`)  
[15b] 2x2 Main effect: Memory errors (H(`r (t15b$pa)`) = `r myround(t15b$s,2)`, *p* `r format_pval(t15b$p.v)`)  
[15c] 2x2 Interaction: Fairness standards x Memory errors (H(`r (t15c$pa)`) = `r myround(t15c$s,2)`, *p* `r format_pval(t15c$p.v)`)  
[15d] Contrast: Discomfort - Upholders  (*W* = `r format(t15d$s,scientific=FALSE)`, *p* `r format_pval(t15d$p)`, *d* = `r myround(t15d$coh,2)`, δ = `r myround(t15d$cl,2)`)   
[15e] Contrast: Discomfort - Violators (*W* = `r format(t15e$s,scientific=FALSE)`, *p* `r format_pval(t15e$p)`, *d* = `r myround(t15e$coh,2)`, δ = `r myround(t15e$cl,2)`)  


```{r s4a_tables,echo=FALSE}

 # apa.reg.table(t9,t10, filename = "s4a_table2.doc")

```

```{r s4a_fig1, fig.height = 4, fig.width = 3.5, fig.asp = 1,echo=FALSE}

#***choose plot type***
  plot = "dist" # to plot figure with full data distribution
  #plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  filename = paste("~/Box Sync/Projects/Active/selfishForgetting/Manuscript/Figures/s4a_fig1_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  pdf(file = filename, #The directory you want to save the file in
  width = 3, # The width of the plot in inches
  height = 3.5) # The height of the plot in inches

#distance specs
  roundUp <- function(x) ceiling(max(x)/10)*10 # this function allows us to round to the nearest 10 above the max data point
  if (plot=="dist") {
    boundary = c(roundUp(max(df4a$GuessVsAve))) # for plotting full distribution plot
    opacity = .15 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = 10 # for plotting zoomed in mean comparison plot
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
  
#pirate plot
  pirateplot(formula = GuessVsAve ~ Fair.split,
             data = df4a,
             pal = "google",
             ylab = "recalled - actual giving (%)",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("dodgerblue1","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, #opacity (.1 = distribution plot, 0 = mean comparison plot)
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("dodgerblue3","firebrick3"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method= inference, 
             inf.disp = "line",
             inf.b.col = c("dodgerblue3","firebrick3"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = 0, #opacity of bar barders
             #bar.b.col = c("darkgrey","darkgrey","darkgrey","darkgrey"),
             bar.f.o = .3,
             bar.f.col = c("dodgerblue1","firebrick1"),
             ylim = c(-boundary, boundary),
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

#ticks & labels
  if (boundary==5|boundary==10) {
  ticks = seq(-boundary,boundary,by=5) # tick positions if boundary is -10 to 10
  } else {
  ticks = seq(-boundary,boundary,by=10) # tick positions if boundary is -max to max
  }
  labels = as.character(ticks) # tick labels
  
#y-axis  
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:2,
       c("upholders", "violators"),
       lwd=0,
       cex.axis=.80,
       pos = min(ticks),
       tck=-0.01) # no ticks on x-axis, just labels

#dashed mid-line
  abline(h = 0, lty = 2,col="gray")

#save output as pdf
   # dev.off()
```

```{r s4a_supp_fig1, fig.height = 4, fig.width = 3.5, fig.asp = 1,echo=FALSE}

#***choose plot type***
  #plot = "dist" # to plot figure with full data distribution
  plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  # filename = paste("~/Box Sync/Projects/Active/selfishForgetting/Manuscript/Figures/s4a_supp_fig1_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  # pdf(file = filename, #The directory you want to save the file in
  # width = 3, # The width of the plot in inches
  # height = 3.5) # The height of the plot in inches

#distance specs
  if (plot=="dist") {
    boundary = round(max(abs(df4a$GuessVsAve))) # for plotting full distribution plot
    opacity = .15 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = 10 # for plotting zoomed in mean comparison plot
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
  
#pirate plot
  pirateplot(formula = GuessVsAveA ~ Fair.split,
             data = df4a,
             pal = "google",
             ylab = "memory errors (absolute %)",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("dodgerblue1","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, #opacity (.1 = distribution plot, 0 = mean comparison plot)
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("dodgerblue3","firebrick3"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method= inference, 
             inf.disp = "line",
             inf.b.col = c("dodgerblue3","firebrick3"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = 0, #opacity of bar barders
             #bar.b.col = c("darkgrey","darkgrey","darkgrey","darkgrey"),
             bar.f.o = .3,
             bar.f.col = c("dodgerblue1","firebrick1"),
             ylim = c(0, boundary),
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

#ticks & labels
  if (boundary==10) {
  ticks = seq(0,boundary,by=5) # tick positions if boundary is -10 to 10
  } else {
  ticks = seq(0,boundary,by=10) # tick positions if boundary is -max to max
  }
  labels = as.character(ticks) # tick labels
  
#y-axis  
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:2,
       label = c("upholders", "violators"),
       lwd=0,
       cex.axis=.80,
       pos = min(ticks),
       tck=-0.01) # no ticks on x-axis, just labels
  
#save output as pdf
   # dev.off()
```

```{r s4a_fig2, fig.height = 4, fig.width = 3.5, fig.asp = 1,echo=FALSE}

#***choose plot type***
  plot = "dist" # to plot figure with full data distribution
  #plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  # filename = paste("~/Box Sync/Projects/Priority/selfishForgetting/Manuscript/Figures/s4a_fig2_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  # pdf(file = filename, #The directory you want to save the file in
  # width = 3, # The width of the plot in inches
  # height = 3.5) # The height of the plot in inches

#distance specs
  if (plot=="dist") {
    boundary = c(1,7) # full range
    opacity = .15 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = c(5,7)
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
       
#pirate plot
  pirateplot(formula = conf ~ Fair.split,
             data = df4a,
             pal = "google",
             ylab = "memory confidence",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("seashell4","dodgerblue1","seashell4","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, #opacity (.1 = distribution plot, 0 = mean comparison plot)
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("dodgerblue3","firebrick3"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method = inference, 
             inf.disp = "line",
             inf.b.col = c("dodgerblue1","firebrick1"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = 0, #opacity of bar barders
             #bar.b.col = c("darkgrey","darkgrey","darkgrey","darkgrey"),
             bar.f.o = .3,
             bar.f.col = c("dodgerblue1","firebrick1"),
             ylim = boundary,
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

#ticks & labels
  ticks = seq(boundary[1],boundary[2],by=1) # tick positions 
  labels = as.character(ticks) # tick labels

#y-axis
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:2,
       label = c("upholders", "violators"),
       lwd=0,
       cex.axis=.8,
       pos = (min(ticks)-.05), # distance between labels and start of y axis
       tck=-0.01) # no ticks on x-axis, just labels

#save output
 # dev.off()
```

```{r s4a_fig3, fig.height = 4, fig.width = 5,fig.asp = .65,echo=FALSE}

# #***choose plot type***
#   #plot = "dist" # to plot figure with full data distribution
#   plot= "means" # to plot figure zoomed in on means
# 
# #setup pdf output 
#   # filename = paste("~/Box Sync/Projects/Priority/selfishForgetting/Manuscript/Figures/s4a_fig3_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
#   # pdf(file = filename, #The directory you want to save the file in
#   # width = 5.625, # The width of the plot in inches
#   # height = 3.5) # The height of the plot in inches
# 
# #setup plot
# 
#   df4a_motmem = select(df4a,motmem,Fair.split,posAdiff,image)
#   df4a_motmem = na.omit(df4a_motmem)
# 
#   #setup plot
#   names(df4a_motmem)[1] = "memory" #rename for plotting
#   df4a_motmem$memory <- as.character(df4a_motmem$memory)
#   df4a_motmem$memory[df4a_motmem$memory == "0"] <- "self-defeating"
#   df4a_motmem$memory[df4a_motmem$memory == "1"] <- "self-serving"
#   df4a_motmem$memory <- as.factor(df4a_motmem$memory)
#   df4a_motmem$memory = factor(df4a_motmem$memory, levels=c("self-defeating","self-serving"))
#   attr(df4a_motmem$Fair.split, "variable.labels") <- "" #so no interaction label appears
#   attr(df4a_motmem$memory, "variable.labels") <- "" #so no interaction label appears
# 
# #distance specs
#   if (plot=="dist") {
#     boundary = c(1,7) # full range
#     opacity = .15 # raw data points visible
#     inference = "ci" # plot confidence interval
#   } else if (plot=="means"){
#     boundary = c(3,7)
#     opacity = 0 # raw data points not visible
#     inference = "se" # plot SEM
#   } else
#     print("specify plot!")
#        
# #pirate plot
#   pirateplot(formula = posAdiff ~ memory + Fair.split,
#              data = df4a_motmem,
#              pal = "google",
#              ylab = "positive affect",
#              yaxt = "n", #suppress y ticks
#              xlab = "",
#              xaxt = "n", #suppress x ticks
#              gl.lty = 0, #rm gridlines
#              bty="n", # remove plot outline
#              cex.axis = .75, #axis size
#              cex.lab = .8, #label size
#              cex.names = .8 , #name size
#              theme = 4,
#              #bean.f.o = .20,
#              #bean.f.col = c("seashell4","dodgerblue1","seashell4","firebrick1"),
#              width.max = .3, #bar width
#              point.o = opacity, #opacity (.1 = distribution plot, 0 = mean comparison plot)
#              point.cex = .5, #size
#              #point.pch = 1,
#              point.col = c("seashell4","dodgerblue3","seashell4","firebrick3"), #color
#              avg.line.o = 0, #avg line
#              #avg.line.col = c("black"),
#              avg.line.lwd = .45,
#              inf.method = inference, 
#              inf.disp = "line",
#              inf.b.col = c("seashell4","dodgerblue1","seashell4","firebrick1"), #color
#              inf.b.o = 1,
#              inf.f.col = c("black","black","black","black"), #color
#              inf.f.o = 1,
#              inf.lwd = .45,
#              bar.b.o = 0, #opacity of bar barders
#              #bar.b.col = c("darkgrey","darkgrey","darkgrey","darkgrey"),
#              bar.f.o = c(.15,.5,.15,.5),
#              bar.f.col = c("dodgerblue1","dodgerblue1","firebrick1","firebrick1"),
#              ylim = boundary,
#              jitter.val = .020) #point jitter
#   
# #tick distance
#   par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels
# 
# #ticks & labels
#   ticks = seq(boundary[1],boundary[2],by=1) # tick positions 
#   labels = as.character(ticks) # tick labels
# 
# #y-axis
#   axis(side = 2,at = ticks,
#         labels=labels,
#         cex.axis=.7, #size of axis tick labels
#         col.ticks = "gray", #color of ticks
#         col="gray", #color of line connecting ticks
#         tck=-0.01,#length of ticks
#         las=1) #orientation of tick labels
# 
# #x-axis
#   axis(side = 1,
#        at = 1:5,
#        label = c("self-defeating","self-serving","","self-defeating","self-serving"),
#        lwd=0,
#        cex.axis=.7,
#        pos = (min(ticks)),
#        tck=-0.01) # no ticks on x-axis, just labels
#   
# title("violators", adj = 0.81, line = -10.5,cex.main=.8,font.main = 1)
# title("upholders", adj = 0.19, line = -10.5,cex.main=.8,font.main = 1)
# 
# ### WARNING: this figure looks bad in R, but the PDF output looks good 8)
# 
# #save output as pdf
#    # dev.off()

```

```{r s4a_respHist,echo=FALSE}

#setup pdf output
  # pdf(file = "~/Box Sync/s4a_respHist.pdf",
  # width = 3,
  # height = 3)

ggplot(data=df4a, aes(df4a$responsible)) +
  geom_histogram(aes(),
                 alpha=.5,
                 bins = 7,
                 position = 'stack',
                 stat = 'bin',
                 color="black",
                 fill="lemon chiffon")+
  theme_classic()+
  xlab("personal responsibility")+ scale_x_continuous(breaks=seq(1,7,1))

#save output
 #dev.off()
```
