---
title: "motMis - study 1"
output: html_document
---
```{r install/load packages,include=FALSE}

#Checks if required packages are installed. Installs them if not.
  # list_pckgs <- c("tidyverse", "psych","lsr","mblm","broom", "effsize","matrixStats","devtools","yarrr","lmtest","lubridate","rstudioapi","broman") # list of required packages
  # new_pckgs <- list_pckgs[!(list_pckgs %in% installed.packages()[,"Package"])] # list of required packages that are not install on this machine
  # if(length(new_pckgs)) install.packages(new_pckgs) # installs them
  
#Load libraries
  library("tidyverse") # plotting, manipulating data etc
  library("psych") # for trait measures
  library("lsr") # for computing effect sizes
  library("mblm") #non-parametric linear regression
  library("broom") # clean outputs
  library("effsize") #non-parametric effect sizes
  library("matrixStats") #calculation SDs for multiple cols
  library("R.matlab") #library to import data from .mat (where it was originally analyzed)
  library("devtools") #pirateplots
  library("yarrr")#pirateplots
  library("sjlabelled")
  library("quantreg")
  library("lme4")
  library("rstudioapi") # load rstudio api
  library("broman") # flexibly display p-values with Rmd function

#Set working directory to the location of this R file (data should also be saved here)
  # curr_dir <- getActiveDocumentContext()$path # get path to currently open R file
  # setwd(dirname(curr_dir)) # set the working directory to curr_dirr
  # #print(getwd()) # print directory
  # rm("curr_dir","new_pckgs","list_pckgs")

```

```{r non-parametric stats functions,echo=FALSE}

#wilcoxon test - paired 
  w.test = function(x,y){
  t = wilcox.test(x,y, paired=TRUE) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x,y) 
    da = cohen.d(x,y)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = da$estimate #add parametric ES
      tbl$cliff.d = d$estimate #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }

#wilcoxon test - one-sample
  w.test_0 = function(x){
  t = wilcox.test(x,mu=0,conf.in=TRUE) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x,0) 
    da = cohensD(x,mu=0)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = da #add parametric ES
      tbl$cliff.d = d$estimate #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }

#mann-whitney test 
  mw.test = function(x,y){
  t = wilcox.test(x~y) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x~y) 
    da = cohen.d(x~y)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = abs(da$estimate) #add parametric ES
      tbl$cliff.d = abs(d$estimate) #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }
    
```

```{r read in data set, include=FALSE}

#read data
  df1 = read.csv("motMis_s1_data.csv") # note: UZH ethics prohibits releasing this dataset, and thus it is not publicly available (for testing, please use the analysis script and data for Experiment 3)

#gender and age
  describe(df1$age)$mean
  table(df1$gender)
 
```

```{r compute variables, include=FALSE}

#compute the true average % transferred
  df1 <- df1 %>% mutate(DGAverage=(DG1+DG2+DG3+DG4+DG5)/5) 
    df1$DGAverage = as.numeric(df1$DGAverage)
    
#compute the % difference between recalled transfer vs. actual transfer
  df1$GuessVsAve = (df1$Guess - df1$DGAverage)

#compute absolute % difference between recalled vs. actual transfer
  df1$GuessVsAveA = abs(df1$GuessVsAve)
  
#compute standard deviation for the five transfer decisions for each subject
  df1_sd <- c('DG1', 'DG2', 'DG3', 'DG4', 'DG5')
    df1 <- df1 %>% mutate(DG.stdev=rowSds(as.matrix(.[df1_sd])))

rm("df1_sd")
 
```

```{r exclusions, echo=FALSE}

  #1 Outliers on our key measure: GuessVsAve

   #ggplot(df1, aes("",GuessVsAve))+geom_boxplot()+theme_bw() #before
    stats = psych::describe(df1$GuessVsAve) #exclude extreme outliers (4 SDs from the mean).
    high_cutoff = stats$mean + (stats$sd*4)
    low_cutoff = stats$mean - (stats$sd*4)
    #SD outliers for memory errors
    df1 <- subset(df1, GuessVsAve < high_cutoff)
    df1 <- subset(df1, GuessVsAve > low_cutoff)
  #ggplot(df1, aes("",GuessVsAve))+geom_boxplot()+theme_bw() #after

rm("stats","high_cutoff","low_cutoff")


#SUPPLEMENTAL EXCLUSIONS:
  #Exclude people who never transferred a positive amount
    # df1 <- subset(df1, DGAverage > 0)
    # df1_0 <- subset(df1, DGAverage == 0)

  
```

```{r grouping,echo=FALSE}

#GIVERS vs. NON-GIVERS: did they give a positive amount?

  df1$giver = as.numeric(df1$DGAverage > 0)

#SELFISH vs. GENEROUS: median split participants based on their average transfers

  m = median(df1$DGAverage) 
  df1_selfish <- subset(df1, DGAverage < m) 
  df1_generous <- subset(df1, DGAverage >= m) 
  
  df1$DGAverage.split = NA 
  df1$DGAverage.split[df1$DGAverage < m] <- "selfish" 
  df1$DGAverage.split[df1$DGAverage >= m] <- "generous" 
  df1$DGAverage.split <- as.factor(df1$DGAverage.split)
  
```

```{r sanity checks, warning=FALSE, message=FALSE, echo=FALSE,include=FALSE}

#[a] are memory errors normally distributed, visually? 
  #ggplot(df1,aes(GuessVsAve))+geom_histogram(binwidth=2)

#[b] are memory errors normally distributed, statistically? 
  shapiro.test(df1$GuessVsAve) 

#[c] can memory errors be explained by noise in DG offers?
  cor.test(df1$DG.stdev,df1$GuessVsAveA, method="spearman")

#[d] can memory errors be explained by RT (attention) to DG offers?
  cor.test(df1$AveRT,df1$GuessVsAveA, method="spearman") 
    
#[e] can memory errors be explained by non-giving?
  cor.test(df1$giver,df1$GuessVsAveA, method="spearman") 
    
#[f] can memory errors be explained by numeracy?
  cor.test(df1$Numeracy,df1$GuessVsAveA, method="spearman") 
    
# group differences
  mw.test(df1$DG.stdev,df1$DGAverage.split)
  mw.test(df1$AveRT,df1$DGAverage.split)
  mw.test(df1$giver,df1$DGAverage.split)
  mw.test(df1$Numeracy,df1$DGAverage.split)
  
```

```{r tests,echo=FALSE,warning=FALSE,include=FALSE}

#[Check normality]
  norm = shapiro.test(df1$GuessVsAve)

#[Descriptives]
  df1_d = describe(df1$DGAverage)
  df1_s_d = describe(df1_selfish$DGAverage)
  df1_p_d = describe(df1_generous$DGAverage)

#[1] Motivated misremembering (main effect)
  t1 = w.test_0(df1$GuessVsAve)

#[2] Motivated misremembering: SELFISH group
  t2 = w.test_0(df1_selfish$GuessVsAve)

#[3] Motivated misremembering: GENEROUS group
  t3 = w.test_0(df1_generous$GuessVsAve)
  
#[4] Motivated misremembering: group difference
  t4 = mw.test(df1$GuessVsAve,df1$DGAverage.split)
  
#[5] Memory accuracy: group difference
  t5 = mw.test(df1$GuessVsAveA,df1$DGAverage.split)

```

```{r knit functions,echo=FALSE}

#formatting p-values for APA
format_pval <- function(x){
  if (x < .001) return(paste('<', '.001'))
  if (x > .1) return(paste('=', myround(x, 2)))
  paste('=', myround(x, 3))   # 3 = no. of digits to round p value to if p < .001 
}

```
#####Check normality

(*W~shapiro-wilk~* = `r myround(norm$s,2)`, *p* `r format_pval(norm$p)`)

#####Descriptives

N = `r round(df1_d$n,1)`, M = `r round(df1_d$mean,1)`   
Selfish N = `r round(df1_s_d$n,1)`, M = `r round(df1_s_d$mean,1)`   
Generous N = `r round(df1_p_d$n,1)`, M = `r round(df1_p_d$mean,1)` 

#####[1] Motivated misremembering (main effect)  

[1] (*V~wilcoxon.sign-ranked~* = `r format(t1$s,scientific=FALSE)`, *p* `r format_pval(t1$p)`, *d* = `r myround(t1$coh,2)`, Cliff's δ = `r myround(t1$cl,2)`)

#####[2] Motivated misremembering: SELFISH group 

[2] (*V* = `r format(t2$s,scientific=FALSE)`, *p* `r format_pval(t2$p)`, *d* = `r myround(t2$coh,2)`, δ = `r myround(t2$cl,2)`)

#####[3] Motivated misremembering: GENEROUS group 

[3] (*V* = `r format(t3$s,scientific=FALSE)`, *p* `r format_pval(t3$p)`, *d* = `r myround(t3$coh,2)`, δ = `r myround(t3$cl,2)`)

#####[4] Motivated misremembering: group difference  

[4] (*W~wilcoxon.rank-sum~* = `r format(t4$s,scientific=FALSE)`, *p* `r format_pval(t4$p)`, *d* = `r myround(t4$coh,2)`, δ = `r myround(t4$cl,2)`)

#####[5] Memory accuracy: group difference

[5] (*W* = `r format(t5$s,scientific=FALSE)`, *p* `r format_pval(t5$p)`, *d* = `r myround(t5$coh,2)`, δ = `r myround(t5$cl,2)`)


```{r s1_fig1, fig.height = .4, fig.width = 3.5,fig.asp = 1,echo=FALSE}

#***choose plot type***
  #plot = "dist" # to plot figure with full data distribution
  plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  # filename = paste("/Users/carlsonr/Box Sync/Projects/Priority/selfishForgetting/Manuscript/Figures/s1_fig1_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  # pdf(file = filename, #The directory you want to save the file in
  # width = 3, # The width of the plot in inches
  # height = 3.5) # The height of the plot in inches

#distance specs
  roundUp <- function(x) ceiling(max(x)/10)*10 # this function allows us to round to the nearest 10 above the max data point
  if (plot=="dist") {
    boundary = roundUp(max(abs(df1$GuessVsAve))) # for plotting full distribution plot
    opacity = .1 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = 10 # for plotting zoomed in mean comparison plot
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
  
  
#pirate plot
  pirateplot(formula = GuessVsAve ~ DGAverage.split,
             data = df1,
             pal = "google",
             ylab = "recalled - actual giving (%)",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("dodgerblue1","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, 
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("gray50","gray10"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method= inference, 
             inf.disp = "line",
             inf.b.col = c("black","black"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = .0, #opacity of bar barders
             bar.b.col = c("gray25","white"),
             bar.lwd = .5,
             bar.f.o = .3,
             bar.f.col = c("gray75","gray25"),
             ylim = c(-boundary, boundary),
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

#ticks & labels
  if (boundary==10) {
  ticks = seq(-boundary,boundary,by=5) # tick positions if boundary is -10 to 10
  } else {
  ticks = seq(-boundary,boundary,by=20) # tick positions if boundary is -max to max
  }
  labels = as.character(ticks) # tick labels
  
#y-axis  
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:2,
       c("generous", "selfish"),
       lwd=0,
       cex.axis=.8,
       pos = min(ticks),
       tck=-0.01) # no ticks on x-axis, just labels
  
#dashed mid-line
  abline(h = 0, lty = 2,col="gray")
  
#detour: our dashed abline drew over the lower end of our error bar -- and it does not look ideal. we cannot draw abline first when using pirateplot, so we'll redraw the bottom end of the error bar manually)
  
  #solution below -- only enable this when plotting mean bars
    sem = sd(df1_generous$GuessVsAve)/sqrt(length(df1_generous$GuessVsAve))
    lower_sem = mean(df1_generous$GuessVsAve)-sem
    segments(x0=.94,x1=1.06,y0=lower_sem,y1=lower_sem,lwd = .45)

#save output as pdf
  # dev.off()

```

```{r s1_supp_fig1, fig.height = .4, fig.width = 3.5,fig.asp = 1,echo=FALSE}

#***choose plot type***
  #plot = "dist" # to plot figure with full data distribution
  plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  # filename = paste("/Users/carlsonr/Box Sync/Projects/Priority/selfishForgetting/Manuscript/Figures/s1_supp_fig1_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  # pdf(file = filename, #The directory you want to save the file in
  # width = 3, # The width of the plot in inches
  # height = 3.5) # The height of the plot in inches

#distance specs
  roundUp <- function(x) ceiling(max(x)/10)*10 # this function allows us to round to the nearest 10 above the max data point
  if (plot=="dist") {
    boundary = roundUp(max(abs(df1$GuessVsAve))) # for plotting full distribution plot
    opacity = .1 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = 10 # for plotting zoomed in mean comparison plot
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
  
#pirate plot
  pirateplot(formula = GuessVsAveA ~ DGAverage.split,
             data = df1,
             pal = "google",
             ylab = "memory errors (absolute %)",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("dodgerblue1","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, 
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("gray50","gray10"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method= inference, 
             inf.disp = "line",
             inf.b.col = c("white","gray"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = .0, #opacity of bar barders
             bar.b.col = c("gray25","white"),
             bar.lwd = .5,
             bar.f.o = .3,
             bar.f.col = c("gray75","gray25"),
             ylim = c(0, boundary),
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

  
#ticks & labels
  if (boundary==10) {
  ticks = seq(0,boundary,by=5) # tick positions if boundary is -10 to 10
  } else {
  ticks = seq(0,boundary,by=10) # tick positions if boundary is -max to max
  }
  labels = as.character(ticks) # tick labels
  
#y-axis  
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1,
       line= 0) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:2,
       label = c("generous", "selfish"),
       lwd=0,
       cex.axis=.80,
       pos = min(ticks),
       tck=-0.01) # no ticks on x-axis, just labels

#save output as pdf
   # dev.off()

```
