---
title: "motMis - study 3"
output: html_document
---
```{r install/load packages,include=FALSE}

#Checks if required packages are installed. Installs them if not.
  # list_pckgs <- c("tidyverse", "psych","lsr","mblm","broom", "effsize","matrixStats","devtools","yarrr","lmtest","lubridate","rstudioapi","apaTables","broman") # list of required packages
  # new_pckgs <- list_pckgs[!(list_pckgs %in% installed.packages()[,"Package"])] # list of required packages that are not install on this machine
  # if(length(new_pckgs)) install.packages(new_pckgs) # installs them

#Load libraries
  library("tidyverse") # plotting, manipulating data etc
  library("psych") # for trait measures
  library("lsr") # for computing effect sizes
  library("mblm") #non-parametric linear regression
  library("broom") # clean outputs
  library("effsize") #non-parametric effect sizes
  library("matrixStats") #calculation SDs for multiple cols
  library("devtools") #pirateplots
  library("yarrr")#pirateplots
  library("lmtest")
  library("lubridate")
  library("rstudioapi") # load rstudio api
  library("apaTables") # for regression tables
  library("broman") # flexibly display p-values with Rmd function
  library("MBESS") # CIs on Rsquared in tables
  
#Set working directory to the location of this R file (data should also be saved here)
  # curr_dir <- getActiveDocumentContext()$path # get path to currently open R file
  # setwd(dirname(curr_dir)) # set the working directory to curr_dirr
  # #print(getwd()) # print directory
  # rm("curr_dir","new_pckgs","list_pckgs")

```

```{r non-parametric stats functions,include=FALSE}

#wilcoxon test - paired
  w.test = function(x,y){
  t = wilcox.test(x,y, paired=TRUE) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x,y) 
    da = cohen.d(x,y)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = da$estimate #add parametric ES
      tbl$cliff.d = d$estimate #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }

#wilcoxon test - one-sample
  w.test_0 = function(x){
  t = wilcox.test(x,mu=0,conf.in=TRUE) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x,0) 
    da = cohensD(x,mu=0)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = da #add parametric ES
      tbl$cliff.d = d$estimate #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }

#mann-whitney test 
  mw.test = function(x,y){
  t = wilcox.test(x~y) #optional: non-parametric CIs with conf.in=TRUE
    d = cliff.delta(x~y) 
    da = cohen.d(x~y)
      tbl = tidy(t) 
      tbl$method <- tbl$alternative <- tbl$estimate <- NULL #remove strings
      tbl$cohen.d = abs(da$estimate) #add parametric ES
      tbl$cliff.d = abs(d$estimate) #add non-parametric ES
      #tbl <- sapply( tbl, as.numeric ) #convert to numeric
        print(tbl) 
  }
     
```

```{r read in data set, include=FALSE}

#read data
  df3 = read.csv("motMis_s3_data.csv") #note: this dataset will be made available upon acceptance of the MS

#gender and age
  describe(df3$age)$mean
  table(df3$gender)
 
```

```{r compute variables, include=FALSE}

#compute the true average % transferred + average RT
  df3 <- df3 %>% mutate(DGAverage=(dg1+dg2+dg3+dg4+dg5)/5, AveRT=(dg1_rt+dg2_rt+dg3_rt+dg4_rt+dg5_rt)/5) 
    df3$DGAverage = as.numeric(df3$DGAverage)
    
#compute the % difference between recalled transfer vs. actual transfer
  df3$GuessVsAve = (df3$Guess - df3$DGAverage)

#compute absolute % difference between recalled vs. actual transfer
  df3$GuessVsAveA = abs(df3$GuessVsAve)
  
#compute standard deviation for the five transfer decisions for each subject
  df3_sd <- c('dg1', 'dg2', 'dg3', 'dg4', 'dg5')
    df3 <- df3 %>% mutate(DG.stdev=rowSds(as.matrix(.[df3_sd])))

#compute difference score between fairness vs actual transfer 
  df3$FairVsActual = (df3$fairQ - df3$DGAverage)

#compute social desirability composite
  df3 <- df3 %>% mutate(socialD=sd_1+sd_2+sd_3+sd_4+sd_5+sd_6+sd_7+sd_8+sd_9+sd_10) #sd6-10 are coded such that FALSE = 1, TRUE = 0, in accordance with the scale's coding rubric.

#compute self-image
  alpha(matrix(c(df3$moral,df3$goodness),ncol=2))
    df3 <- df3 %>% mutate(image=(moral+goodness)/2)

#compute feelings
  alpha(matrix(c(df3$pos,df3$hap),ncol=2))
    df3 <- df3 %>% mutate(feelings=(pos+hap)/2)

#change time to numeric. 
  df3$studytime <- ms(df3$studytime) #note, lubridate confuses minutes for seconds here 
    df3$studytime <- as.numeric(df3$studytime) #minutes converted to numeric

rm("df3_sd")
 
```

```{r exclusions,include=FALSE} 

#EXCLUSIONS

  #1 Failed to follow instructions: bottest = 0 (failed bottest)

  #count(df3$bottest_filter == 0)
  df3 <- subset(df3, bottest_filter == 1)

  #2 Failed comprehension: deception1 = 0 (instructions NOT clear about money)

  #count(df3$deception1 == 5)
  df3 <- subset(df3, deception1 == 4)

  #3 Failed comprehension: deception2 = 0 (intructions NOT clear about other participant)

  #count(df3$deception2 == 2)
  df3 <- subset(df3, deception2 == 1)

  #4 Reported skepticism about the study: feedbackexcl = 0 (did NOT believe in P2 / or confused)

  #count(df3$feedback_excl == 0)
  df3 <- subset(df3, feedback_excl == 1)

  #5 Outliers on our key measure: GuessVsAve

  #ggplot(df3, aes("",GuessVsAve))+geom_boxplot()+theme_classic() # before
    stats = psych::describe(df3$GuessVsAve) #exclude extreme outliers (4 SDs from the mean).
    high_cutoff = stats$mean + (stats$sd*4)
    low_cutoff = stats$mean - (stats$sd*4)

    #SD outliers for memory errors
    count(df3$GuessVsAve >= high_cutoff)
    df3 <- subset(df3, GuessVsAve < high_cutoff)

    count(df3$GuessVsAve <= low_cutoff)
    df3 <- subset(df3, GuessVsAve > low_cutoff)
  #ggplot(df3, aes("",GuessVsAve))+geom_boxplot()+theme_classic() # after

rm("stats","high_cutoff","low_cutoff")


#SUPPLEMENTAL EXCLUSIONS:
  #Exclude people who never transferred a positive amount
    # df3 <- subset(df3, DGAverage > 0)
    # df3_0 <- subset(df3, DGAverage == 0)


```

```{r grouping, include=FALSE, echo=FALSE}

#GIVERS vs. NON-GIVERS: did they give a positive amount?

  df3$giver = as.numeric(df3$DGAverage > 0)


#SELFISH vs. GENEROUS: median split participants based on their average transfers

  m = median(df3$DGAverage,na.rm = TRUE) 
  df3_selfish <- subset(df3, DGAverage < m) 
  df3_generous <- subset(df3, DGAverage >= m) 
  
  df3$DGAverage.split = NA 
  df3$DGAverage.split[df3$DGAverage < m] <- "selfish" 
  df3$DGAverage.split[df3$DGAverage >= m] <- "generous" 
  df3$DGAverage.split <- as.factor(df3$DGAverage.split)
    
  
#NORM VIOLATORS vs. NORM UPHOLDERS: split by whether people upheld or violated their own fairness norms
    
  df3_fairness_violated <- subset(df3, DGAverage < fairQ) # gave less than norm
  df3_fairness_upheld <- subset(df3, DGAverage >= fairQ) # gave equal or more than norm

  df3$Fair.split <- NA
  df3$Fair.split[df3$DGAverage < df3$fairQ] <- "violators"
  df3$Fair.split[df3$DGAverage >= df3$fairQ] <- "upholders"
  df3$Fair.split <- as.factor(df3$Fair.split)

    
#UNGENEROUS vs. GENEROUS norm upholders: split on generosity within fairness upholders
  
  m_fair = median(df3_fairness_upheld$DGAverage,na.rm = TRUE) 
  df3_fair_selfish <- subset(df3_fairness_upheld, DGAverage < m_fair)
  df3_fair_generous <- subset(df3_fairness_upheld, DGAverage >= m_fair)
  
  df3_fairness_upheld$DGAverage.split = NA 
  df3_fairness_upheld$DGAverage.split[df3_fairness_upheld$DGAverage < m_fair] <- "selfish"
  df3_fairness_upheld$DGAverage.split[df3_fairness_upheld$DGAverage >= m_fair] <- "generous"  
  df3_fairness_upheld$DGAverage.split <- as.factor(df3_fairness_upheld$DGAverage.split)


#MOTIVATED MEMORY: create variables capturing positive (motivated) or negative (unmotivated) memory errors
  
  df3$motmem = NA
  df3$motmem[df3$GuessVsAve < 0] = 0
  df3$motmem[df3$GuessVsAve > 0] = 1
  df3$motmem = as.factor(df3$motmem)
  
  df3_fairness_upheld$motmem = NA
  df3_fairness_upheld$motmem[df3_fairness_upheld$GuessVsAve < 0] = 0
  df3_fairness_upheld$motmem[df3_fairness_upheld$GuessVsAve > 0] = 1
  df3_fairness_upheld$motmem = as.factor(df3_fairness_upheld$motmem)
  
  df3_fairness_violated$motmem = NA
  df3_fairness_violated$motmem[df3_fairness_violated$GuessVsAve < 0] = 0
  df3_fairness_violated$motmem[df3_fairness_violated$GuessVsAve > 0] = 1
  df3_fairness_violated$motmem = as.factor(df3_fairness_violated$motmem)

```

```{r sanity checks, warning=FALSE, message=FALSE, echo=FALSE,include=FALSE}

#[a] are memory errors non-normally distributed, visually? 
  #ggplot(df3,aes(GuessVsAve))+geom_histogram(binwidth=2)

#[b] are memory errors non-normally distributed, statistically? 
  shapiro.test(df3$GuessVsAve) 

#[c] can memory errors be explained by noise in DG offers?
  cor.test(df3$DG.stdev,df3$GuessVsAveA, method="spearman") 

#[d] can memory errors be explained by RT (attention) to DG offers?
  cor.test(df3$AveRT,df3$GuessVsAveA, method="spearman") 

#[e] can memory errors be explained by non-giving?
  cor.test(df3$giver,df3$GuessVsAveA, method="spearman") 

#[e] can memory errors be explained by how long people took to complete the study?
  cor.test(df3$studytime,df3$GuessVsAveA,method="spearman")

#[f] can memory errors be explained by numeracy?
  cor.test(df3$Numeracy,df3$GuessVsAveA, method="spearman") 
  
#[g] can memory errors be explained by social desirability?
  cor.test(df3$socialD,df3$GuessVsAveA, method="spearman")
  
#######

#[h] internal validity of affective state measure
  a_validity = df3[c("pos","hap")]	
  psych::alpha(a_validity)
  
#[i] internal validity of moral self-view measure
  m_validity = df3[c("goodness","moral")]	
  psych::alpha(m_validity)
  
# group differences
  
  mw.test(df3$DG.stdev,df3$Fair.split)
  mw.test(df3$AveRT,df3$Fair.split)
  mw.test(df3$giver,df3$Fair.split)
  mw.test(df3$Numeracy,df3$Fair.split)
  mw.test(df3$socialD,df3$Fair.split)
  
```

```{r tests,echo=FALSE,warning=FALSE,include=FALSE}

#[Check normality]
  norm = shapiro.test(df3$GuessVsAve)

#[Descriptives]
  df3_d = describe(df3$DGAverage)
  df3_v_d = describe(df3_fairness_violated$DGAverage)
  df3_u_d = describe(df3_fairness_upheld$DGAverage)

#[1] Motivated misremembering (main effect)
  t1 = w.test_0(df3$GuessVsAve)

#[2] Motivated misremembering: VIOLATOR group
  t2 = w.test_0(df3_fairness_violated$GuessVsAve)

#[3] Motivated misremembering: UPHOLDER group
  t3 = w.test_0(df3_fairness_upheld$GuessVsAve)
  
#[4] Motivated misremembering: group difference
  t4 = mw.test(df3$GuessVsAve,df3$Fair.split)
  
#[5] Memory accuracy: group difference
  t5 = mw.test(df3$GuessVsAveA,df3$Fair.split)
  
##################################################
  
#[Descriptives]
  df3_fu_d = describe(df3_fair_selfish$DGAverage)
  df3_fg_d = describe(df3_fair_generous$DGAverage)
  
#[6] Motivated misremembering: fair UNGENEROUS group
  t6 = w.test_0(df3_fair_selfish$GuessVsAve)

#[7] Motivated misremembering: fair GENEROUS group
  t7 = w.test_0(df3_fair_generous$GuessVsAve)
  
#[8] Motivated misremembering: group difference
  t8 = mw.test(df3_fairness_upheld$GuessVsAve,df3_fairness_upheld$DGAverage.split)

##################################################

#[9] Model predicting recalled generosity from actual geneorsity (controlling for decision speed + variance + non-giving + numeracy)
    t9 <- lm(Guess ~ DGAverage + AveRT + DG.stdev + giver + Numeracy + socialD, data=df3)
    tidy(t9)
    t9a = AIC(t9)
    t9b = BIC(t9)
    t9l = logLik(t9)
      t9l
      
#[10] Model predicting recalled generosity from actual geneorsity fairness deviations (controlling for decision speed + variance + non-giving + numeracy)
    t10 <- lm(Guess ~ DGAverage + AveRT + DG.stdev + giver + Numeracy + socialD + FairVsActual, data=df3)
    tidy(t10)
    t10a = AIC(t10)
    t10b = BIC(t10)
    t10l = logLik(t10)
      t10l

#[11] Comparison of models' goodness of fit
    t11 = lrtest(t9,t10)
    t11

##################################################

#[12] Confidence: Correlations
  t12a = cor.test(df3$conf,df3$GuessVsAve, method="spearman") 
  t12b = cor.test(df3$conf,df3$GuessVsAveA, method="spearman") 
  
#[13] Confidence: Group difference
  t13 = mw.test(df3$conf,df3$Fair.split)
  
#[14] Moral and emotional benefits
  t14a = mw.test(df3$image,df3$Fair.split)
  t14b = mw.test(df3$feelings,df3$Fair.split)
  
#[15] Moral benefits of memory errors: Upholders & Violators
  t15a = kruskal.test(image ~ Fair.split, data = df3)
  t15b = kruskal.test(image ~ motmem, data = df3)
  int = interaction(df3$Fair.split, df3$motmem)
  t15c = kruskal.test(image ~ int, data = df3)  
  t15d = mw.test(df3_fairness_upheld$image,df3_fairness_upheld$motmem)
  t15e = mw.test(df3_fairness_violated$image,df3_fairness_violated$motmem)

#[16] Emotional benefits of memory errors: Upholders & Violators
  t16a = kruskal.test(feelings ~ Fair.split, data = df3)
  t16b = kruskal.test(feelings ~ motmem, data = df3)
  int = interaction(df3$Fair.split, df3$motmem)
  t16c = kruskal.test(feelings ~ int, data = df3)  
  t16d = mw.test(df3_fairness_upheld$feelings,df3_fairness_upheld$motmem)
  t16e = mw.test(df3_fairness_violated$feelings,df3_fairness_violated$motmem)
  
```

```{r knit functions,echo=FALSE}

#formatting p-values for APA
format_pval <- function(x){
  if (x < .001) return(paste('<', '.001'))
  if (x > .1) return(paste('=', myround(x, 2)))
  paste('=', myround(x, 3))   # 3 = no. of digits to round p value to if p < .001 
}

```
#####Check normality

(*W* = `r myround(norm$s,2)`, *p* `r format_pval(norm$p)`)

#####Descriptives

N = `r round(df3_d$n,1)`, M = `r myround(df3_d$mean,1)`   
Violators N = `r round(df3_v_d$n,1)`, M = `r myround(df3_v_d$mean,1)`   
Upholders N = `r round(df3_u_d$n,1)`, M = `r myround(df3_u_d$mean,1)` 

#####[1] Motivated misremembering (main effect)  

[1] (*V* = `r format(t1$s,scientific=FALSE)`, *p* `r format_pval(t1$p)`, *d* = `r myround(t1$coh,2)`, δ = `r myround(t1$cl,2)`)

#####[2] Motivated misremembering: VIOLATORS group 

[2] (*V* = `r format(t2$s,scientific=FALSE)`, *p* `r format_pval(t2$p)`, *d* = `r myround(t2$coh,2)`, δ = `r myround(t2$cl,2)`)

#####[3] Motivated misremembering: UPHOLDERS group 

[3] (*V* = `r format(t3$s,scientific=FALSE)`, *p* `r format_pval(t3$p)`, *d* = `r myround(t3$coh,2)`, δ = `r myround(t3$cl,2)`)

#####[4] Motivated misremembering: group difference  

[4] (*W* = `r format(t4$s,scientific=FALSE)`, *p* `r format_pval(t4$p)`, *d* = `r myround(t4$coh,2)`, δ = `r myround(t4$cl,2)`)

#####[5] Memory accuracy: group difference  

[5] (*W* = `r format(t5$s,scientific=FALSE)`, *p* `r format_pval(t5$p)`, *d* = `r myround(t5$coh,2)`, δ = `r myround(t5$cl,2)`)

#####Descriptives

N = `r round(df3_d$n,1)`, M = `r myround(df3_d$mean,1)`   
Fair ungenerous N = `r round(df3_fu_d$n,1)`, M = `r myround(df3_fu_d$mean,1)`   
Fair generous N = `r round(df3_fg_d$n,1)`, M = `r myround(df3_fg_d$mean,1)`

#####[6] Motivated misremembering: FAIR ungenerous group 

[6] (*V* = `r format(t6$s,scientific=FALSE)`, *p* `r format_pval(t6$p)`, *d* = `r myround(t6$coh,2)`, δ = `r myround(t6$cl,2)`)

#####[7] Motivated misremembering: FAIR generous group 

[7] (*V* = `r format(t7$s,scientific=FALSE)`, *p* `r format_pval(t7$p)`, *d* = `r myround(t7$coh,2)`, δ = `r myround(t7$cl,2)`)

#####[8] Motivated misremembering: group difference  

[8] (*W* = `r format(t8$s,scientific=FALSE)`, *p* `r format_pval(t8$p)`, *d* = `r myround(t8$coh,2)`, δ = `r myround(t8$cl,2)`)

#####[9-11] Model comparison: recall from actual generosity (i) or actual generosity + fairness deviations (ii)

[9-11] (ΔAIC = `r myround(t9a-t10a,2)`, ΔBIC = `r myround(t9b-t10b,2)`, χ2 (`r (t11$D[2])`) = `r myround(t11$Ch[2],2)`, *p* `r format_pval(t11$P[2])`)

#####[12] Confidence: Correlations

[12a]  Signed memory errors (*r~s~* = `r myround(t12a$e,2)`, *p* `r format_pval(t12a$p.v)`)  
[12b]  Unsigned memory errors (*r~s~* = `r myround(t12b$e,2)`, *p* `r format_pval(t12b$p.v)`)

#####[13] Confidence: Group difference
[13] (*W* = `r format(t13$s,scientific=FALSE)`, *p* `r format_pval(t13$p)`, *d* = `r myround(t13$coh,2)`, δ = `r myround(t13$cl,2)`)

#####[14] Moral and emotional benefits
[14a] Moral image (*W* = `r format(t14a$s,scientific=FALSE)`, *p* `r format_pval(t14a$p)`, *d* = `r myround(t14a$coh,2)`, δ = `r myround(t14a$cl,2)`)   
[14b] Positive emotion (*W* = `r format(t14b$s,scientific=FALSE)`, *p* `r format_pval(t14b$p)`, *d* = `r myround(t14b$coh,2)`, δ = `r myround(t14b$cl,2)`)

#####[15] Moral benefits of memory errors: Upholders & Violators

[15a] 2x2 Main effect: Fairness Standards (*H~kruskal.wallis~*(`r (t15a$pa)`) = `r myround(t15a$s,2)`, *p* `r format_pval(t15a$p.v)`)  
[15b] 2x2 Main effect: Memory errors (H(`r (t15b$pa)`) = `r myround(t15b$s,2)`, *p* `r format_pval(t15b$p.v)`)  
[15c] 2x2 Interaction: Fairness standards x Memory errors (H(`r (t15c$pa)`) = `r myround(t15c$s,2)`, *p* `r format_pval(t15c$p.v)`)  
[15d] Contrast: Moral image - Upholders  (*W* = `r format(t15d$s,scientific=FALSE)`, *p* `r format_pval(t15d$p)`, *d* = `r myround(t15d$coh,2)`, δ = `r myround(t15d$cl,2)`)   
[15e] Contrast: Moral image - Violators (*W* = `r format(t15e$s,scientific=FALSE)`, *p* `r format_pval(t15e$p)`, *d* = `r myround(t15e$coh,2)`, δ = `r myround(t15e$cl,2)`)  


#####[16] Emotional benefits of memory errors: Upholders & Violators

[16a] 2x2 Main effect: Fairness Standards (H(`r (t16a$pa)`) = `r myround(t16a$s,2)`, *p* `r format_pval(t16a$p.v)`)  
[16b] 2x2 Main effect: Memory errors (H(`r (t16b$pa)`) = `r myround(t16b$s,2)`, *p* `r format_pval(t16b$p.v)`)  
[16c] 2x2 Interaction: Fairness standards x Memory errors (H(`r (t16c$pa)`) = `r myround(t16c$s,2)`, *p* `r format_pval(t16c$p.v)`)    
[16d] Contrast: Positive emotion - Upholders (*W* = `r format(t16d$s,scientific=FALSE)`, *p* `r format_pval(t16d$p)`, *d* = `r myround(t16d$coh,2)`, δ = `r myround(t16d$cl,2)`)   
[16e] Contrast: Positive emotion - Violators (*W* = `r format(t16e$s,scientific=FALSE)`, *p* `r format_pval(t16e$p)`, *d* = `r myround(t16e$coh,2)`, δ = `r myround(t16e$cl,2)`)

```{r s3_tables,echo=FALSE}

  #apa.reg.table(t9,t10, filename = "s3_table2.doc")

```

```{r s3_fig1, fig.height = 4, fig.width = 3.5, fig.asp = 1,echo=FALSE}

#***choose plot type***
  plot = "dist" # to plot figure with full data distribution
  #plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  # filename = paste("/Users/carlsonr/Box Sync/Projects/Priority/selfishForgetting/Manuscript/Figures/s3_fig1_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  # pdf(file = filename, #The directory you want to save the file in
  # width = 3, # The width of the plot in inches
  # height = 3.5) # The height of the plot in inches

#distance specs
  roundUp <- function(x) ceiling(max(x)/10)*10 # this function allows us to round to the nearest 10 above the max data point
  if (plot=="dist") {
    boundary = c(roundUp(max(df3$GuessVsAve))) # for plotting full distribution plot
    opacity = .1 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = 10 # for plotting zoomed in mean comparison plot
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
  
#pirate plot
  pirateplot(formula = GuessVsAve ~ Fair.split,
             data = df3,
             pal = "google",
             ylab = "recalled - actual giving (%)",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("dodgerblue1","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, #opacity (.1 = distribution plot, 0 = mean comparison plot)
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("dodgerblue3","firebrick3"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method= inference, 
             inf.disp = "line",
             inf.b.col = c("dodgerblue3","firebrick3"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = 0, #opacity of bar barders
             #bar.b.col = c("darkgrey","darkgrey","darkgrey","darkgrey"),
             bar.f.o = .3,
             bar.f.col = c("dodgerblue1","firebrick1"),
             ylim = c(-boundary, boundary),
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

#ticks & labels
  if (boundary==10) {
  ticks = seq(-boundary,boundary,by=5) # tick positions if boundary is -10 to 10
  } else {
  ticks = seq(-boundary,boundary,by=10) # tick positions if boundary is -max to max
  }
  labels = as.character(ticks) # tick labels
  
#y-axis  
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:2,
       c("upholders", "violators"),
       lwd=0,
       cex.axis=.80,
       pos = min(ticks),
       tck=-0.01) # no ticks on x-axis, just labels

#dashed mid-line
  abline(h = 0, lty = 2,col="gray")

#save output as pdf
   # dev.off()
```

```{r s3_supp_fig1, fig.height = 4, fig.width = 3.5, fig.asp = 1,echo=FALSE}

#***choose plot type***
  plot = "dist" # to plot figure with full data distribution
  #plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  # filename = paste("/Users/carlsonr/Box Sync/Projects/Priority/selfishForgetting/Manuscript/Figures/s3_supp_fig1_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  # pdf(file = filename, #The directory you want to save the file in
  # width = 3, # The width of the plot in inches
  # height = 3.5) # The height of the plot in inches

#distance specs
  if (plot=="dist") {
    boundary = round(max(abs(df3$GuessVsAve))) # for plotting full distribution plot
    opacity = .1 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = 10 # for plotting zoomed in mean comparison plot
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
  
#pirate plot
  pirateplot(formula = GuessVsAveA ~ Fair.split,
             data = df3,
             pal = "google",
             ylab = "memory errors (absolute %)",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("dodgerblue1","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, #opacity (.1 = distribution plot, 0 = mean comparison plot)
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("dodgerblue3","firebrick3"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method= inference, 
             inf.disp = "line",
             inf.b.col = c("dodgerblue3","firebrick3"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = 0, #opacity of bar barders
             #bar.b.col = c("darkgrey","darkgrey","darkgrey","darkgrey"),
             bar.f.o = .3,
             bar.f.col = c("dodgerblue1","firebrick1"),
             ylim = c(0, boundary),
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

#ticks & labels
  if (boundary==10) {
  ticks = seq(0,boundary,by=5) # tick positions if boundary is -10 to 10
  } else {
  ticks = seq(0,boundary,by=10) # tick positions if boundary is -max to max
  }
  labels = as.character(ticks) # tick labels
  
#y-axis  
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:2,
       label = c("upholders", "violators"),
       lwd=0,
       cex.axis=.80,
       pos = min(ticks),
       tck=-0.01) # no ticks on x-axis, just labels
  
#save output as pdf
   # dev.off()
```

```{r s3_fig2, fig.height = 4, fig.width = 3.5, fig.asp = 1,echo=FALSE}

#***choose plot type***
  plot = "dist" # to plot figure with full data distribution
  #plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  # filename = paste("/Users/carlsonr/Box Sync/Projects/Priority/selfishForgetting/Manuscript/Figures/s3_fig2_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  # pdf(file = filename, #The directory you want to save the file in
  # width = 3, # The width of the plot in inches
  # height = 3.5) # The height of the plot in inches

#distance specs
  if (plot=="dist") {
    boundary = c(1,7) # full range
    opacity = .1 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = c(5,7)
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
       
#pirate plot
  pirateplot(formula = conf ~ Fair.split,
             data = df3,
             pal = "google",
             ylab = "memory confidence",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("seashell4","dodgerblue1","seashell4","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, #opacity (.1 = distribution plot, 0 = mean comparison plot)
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("dodgerblue3","firebrick3"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method = inference, 
             inf.disp = "line",
             inf.b.col = c("dodgerblue1","firebrick1"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = 0, #opacity of bar barders
             #bar.b.col = c("darkgrey","darkgrey","darkgrey","darkgrey"),
             bar.f.o = .3,
             bar.f.col = c("dodgerblue1","firebrick1"),
             ylim = boundary,
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

#ticks & labels
  ticks = seq(boundary[1],boundary[2],by=1) # tick positions 
  labels = as.character(ticks) # tick labels

#y-axis
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:2,
       label = c("upholders", "violators"),
       lwd=0,
       cex.axis=.8,
       pos = (min(ticks)-.05), # distance between labels and start of y axis
       tck=-0.01) # no ticks on x-axis, just labels

#save output
 # dev.off()
```

```{r s3_fig3, fig.height = 4, fig.width = 5,fig.asp = .65,echo=FALSE}

#***choose plot type***
  #plot = "dist" # to plot figure with full data distribution
  plot= "means" # to plot figure zoomed in on means

#setup pdf output 
  # filename = paste("/Users/carlsonr/Box Sync/Projects/Priority/selfishForgetting/Manuscript/Figures/s3_fig3_", plot,".pdf", sep="") #Saves file with 'dist' or 'means' name extension
  # pdf(file = filename, #The directory you want to save the file in
  # width = 5.625, # The width of the plot in inches
  # height = 3.5) # The height of the plot in inches

#setup plot

  df3_motmem = select(df3,motmem,Fair.split,feelings,image)
  df3_motmem = na.omit(df3_motmem)

  #setup plot
  names(df3_motmem)[1] = "memory" #rename for plotting
  df3_motmem$memory <- as.character(df3_motmem$memory)
  df3_motmem$memory[df3_motmem$memory == "0"] <- "self-defeating"
  df3_motmem$memory[df3_motmem$memory == "1"] <- "self-serving"
  df3_motmem$memory <- as.factor(df3_motmem$memory)
  df3_motmem$memory = factor(df3_motmem$memory, levels=c("self-defeating","self-serving"))
  attr(df3_motmem$Fair.split, "variable.labels") <- "" #so no interaction label appears
  attr(df3_motmem$memory, "variable.labels") <- "" #so no interaction label appears

#distance specs
  if (plot=="dist") {
    boundary = c(1,7) # full range
    opacity = .1 # raw data points visible
    inference = "ci" # plot confidence interval
  } else if (plot=="means"){
    boundary = c(3,7)
    opacity = 0 # raw data points not visible
    inference = "se" # plot SEM
  } else
    print("specify plot!")
       
#pirate plot
  pirateplot(formula = feelings ~ memory + Fair.split,
             data = df3_motmem,
             pal = "google",
             ylab = "positive affect",
             yaxt = "n", #suppress y ticks
             xlab = "",
             xaxt = "n", #suppress x ticks
             gl.lty = 0, #rm gridlines
             bty="n", # remove plot outline
             cex.axis = .75, #axis size
             cex.lab = .8, #label size
             cex.names = .8 , #name size
             theme = 4,
             #bean.f.o = .20,
             #bean.f.col = c("seashell4","dodgerblue1","seashell4","firebrick1"),
             width.max = .3, #bar width
             point.o = opacity, #opacity (.1 = distribution plot, 0 = mean comparison plot)
             point.cex = .5, #size
             #point.pch = 1,
             point.col = c("seashell4","dodgerblue3","seashell4","firebrick3"), #color
             avg.line.o = 0, #avg line
             #avg.line.col = c("black"),
             avg.line.lwd = .45,
             inf.method = inference, 
             inf.disp = "line",
             inf.b.col = c("seashell4","dodgerblue1","seashell4","firebrick1"), #color
             inf.b.o = 1,
             inf.f.col = c("black","black","black","black"), #color
             inf.f.o = 1,
             inf.lwd = .45,
             bar.b.o = 0, #opacity of bar barders
             #bar.b.col = c("darkgrey","darkgrey","darkgrey","darkgrey"),
             bar.f.o = c(.15,.5,.15,.5),
             bar.f.col = c("dodgerblue1","dodgerblue1","firebrick1","firebrick1"),
             ylim = boundary,
             jitter.val = .020) #point jitter
  
#tick distance
  par(mgp = c(2, .5, 0)) # middle value changes distance between ticks and tick labels

#ticks & labels
  ticks = seq(boundary[1],boundary[2],by=1) # tick positions 
  labels = as.character(ticks) # tick labels

#y-axis
  axis(side = 2,at = ticks,
        labels=labels,
        cex.axis=.7, #size of axis tick labels
        col.ticks = "gray", #color of ticks
        col="gray", #color of line connecting ticks
        tck=-0.01,#length of ticks
        las=1) #orientation of tick labels

#x-axis
  axis(side = 1,
       at = 1:5,
       label = c("self-defeating","self-serving","","self-defeating","self-serving"),
       lwd=0,
       cex.axis=.7,
       pos = (min(ticks)),
       tck=-0.01) # no ticks on x-axis, just labels
  
title("violators", adj = 0.81, line = -10.5,cex.main=.8,font.main = 1)
title("upholders", adj = 0.19, line = -10.5,cex.main=.8,font.main = 1)

### WARNING: this figure looks bad in R, but the PDF output looks good 8)

#save output as pdf
   # dev.off()
```
